---
title: 'Reproducible Research: Peer Assessment 1'
output:html_document:
keep_md: yes
---

# Reproducible Research: Peer Assessment 1

SETDIGIT AND LOCALE
Created February 2016

## Loading and preprocessing the data

**Note:** It is required that you download the [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r, Loading and processing data, echo = TRUE}

## Unzip activity.zip if needed
if( !file.exists("activity.csv") ) {
  temp <- tempfile()
  file <- unzip("activity.zip")
  unlink(temp)
}

## Read data
activity <- read.csv("activity.csv")

## convert date variable to dates
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

## Mean total number of steps taken per day

```{r, Number steps per day, echo = TRUE }

SumStepsPerDay<- with( activity, aggregate( steps ~ date, FUN = sum))

```

### Histogram of total number of steps per day
```{r histogram, echo = TRUE, fig.height=4}

## histogram of the total number of steps taken each day
library(ggplot2)
g <- ggplot(data = SumStepsPerDay, aes(x = steps)) + geom_histogram(stat  ="bin", fill = "steelblue", binwidth = 1000) + ggtitle("Number of steps taken each day" ) + xlab("Steps per day")
print(g)
#png( file = "./figure/Steps_per_Day_Histogram.png", height = 480, width = 480, unit = "px")
#print(p)
#dev.off()

```

```{r, mean and median, echo =TRUE }
## Calculates Mean (rounded) and median of steps taken each day

MeanStepsPerDay <- round( mean(SumStepsPerDay$steps), digit = 2 )
print(MeanStepsPerDay)

MedianStepsPerDay <- median(SumStepsPerDay$steps)
print(MedianStepsPerDay)

```

The mean and median of steps taken per day are `r MeanStepsPerDay` and `r MedianStepsPerDay` respectively.

##  Average daily activity pattern

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, Time series plot, echo = TRUE, fig.height=4}

## Average number of steps per 5 minutes interval
MeanStepsPerInterval <- with( activity, aggregate( steps ~ interval, FUN = mean, na.rm = TRUE ))

q <- qplot( x = interval, y = steps, data = MeanStepsPerInterval, geom = "line", main = "Average number of steps per 5 minutes intervals")
print(q)

MaxInterval <- MeanStepsPerInterval$interval[ which.max( MeanStepsPerInterval$steps  ) ] 
MaxSteps <- MeanStepsPerInterval$steps[ which.max( MeanStepsPerInterval$steps  ) ] 
```
The time interval containing the maximum number of steps is `r MaxInterval` with an average of `r MaxSteps`


## Imputing missing values

### Number of lines with NA values

```{r, count NAs, echo = TRUE }
naNum <- nrow( activity[is.na(activity$steps), ] )
print(naNum)

```
There are `r naNum` rows with missing values in the activity dataset

### Imputing missing data

```{r, Explore missing data, echo = TRUE }
MissingDays <- unique(activity$date[is.na(activity$steps)])
print(MissingDays)
MissingDaysMean<- with( activity[ activity$date %in% MissingDays, ], tapply( steps, date, sum, na.rm = TRUE) )
print(MissingDaysMean)
```
There are 8 days that contain no data at all, while the others appear to be complete.

For each of those, each time interval will be replaced by the mean number of steps for this particular time interval over the entire dataset

**Note** This will add a new column (ImputedSteps) to the activity dataset rather than replacing the steps variable

```{r, impute missing data, echo = TRUE}

## Create a vector with imputed step values if required
ImputedSteps <- NULL
ImputedValue <- NULL
for( i in 1:nrow(activity) ){
    if( is.na(activity$steps[i]) ) {
        MissingInterval <- activity$interval[i]
        ImputedValue <- MeanStepsPerInterval$steps[MeanStepsPerInterval$interval==MissingInterval]
    } else { ImputedValue <- activity$steps[i] }
    ImputedSteps <- c(ImputedSteps, ImputedValue)
}

## Add Imputed Steps values to the activity dataset
activity$ImputedSteps <- ImputedSteps

```

#### The new histogram with imputed values

```{r, New Histogram, echo = TRUE, fig.height=4 }

## Calculate sum of steps per days
NewSumStepsPerDay<- with( activity, aggregate( ImputedSteps ~ date, FUN = sum))

## Creates histogram
g <- ggplot(data = NewSumStepsPerDay, aes(x = ImputedSteps)) + geom_histogram(stat  ="bin", fill = "magenta", binwidth = 1000) + ggtitle("Number of steps taken each day (Imputed data)" ) + xlab("Steps per day")
print(g)
```

#### Comparing mean and median values

```{r, New Mean and Median}
NewMeanStepsPerDay <- round( mean(NewSumStepsPerDay$ImputedSteps), digit = 2 )
print(NewMeanStepsPerDay)
NewMedianStepsPerDay <- round( median(NewSumStepsPerDay$ImputedSteps), digit = 2)
print(NewMedianStepsPerDay)

```

The mean of steps taken per day is unchanged (`r MeanStepsPerDay` with NA and `r NewMeanStepsPerDay` with imputed data. The median is changes from `r MedianStepsPerDay` with NA to `r NewMedianStepsPerDay`.

## Differences in activity patterns between weekdays and weekends

Create a daytype factor variable with two levels, "weekday" and "weekend"
```{r, Set a weekday/weekend factor, echo = TRUE }

activity$weekday <- weekdays(as.Date(activity$date))
WeekendsDay <- c("Saturday", "Sunday")
activity$daytype <- as.factor( ifelse(activity$weekday %in% WeekendsDay, "weekend", "weekday") )

```

Panel Plot
```{r, Panel plot, echo = TRUE}

## Calculate mean steps per interval
MeanStepsPerIntervalWeek <- with( activity, aggregate( ImputedSteps ~ interval + daytype, FUN = mean, na.rm = TRUE ))

## Panel plot
g <- ggplot(data = MeanStepsPerIntervalWeek , aes(x = interval, y = ImputedSteps, colour = daytype)) + geom_line(stat  ="identity") + ggtitle("Mean steps per 5 minutes interval in weekdays and weekends" ) + facet_grid(.~daytype)
print(g)
```

The mean number of steps at early time intervals appears to be higher in weekdays than for weekends