---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, unzip and read data convert to date, echo = TRUE}
## Unzip activity.zip if needed
if( !file.exists("activity.csv") ) {
  temp <- tempfile()
  file <- unzip("activity.zip")
  unlink(temp)
}

## Read data
activity <- read.csv("activity.csv")

## convert date variable to dates
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE }

## Total number of steps taken per day
SumStepsPerDay <- with( activity, tapply( steps, date, sum, na.rm = TRUE ))
```

Histogram of total number of steps per day
```{r, echo = TRUE, fig.height=4}
## histogram of the total number of steps taken each day
library(ggplot2)
p <- qplot( SumStepsPerDay) + geom_vline( xintercept = mean(SumStepsPerDay), col = "red", size = 2) + geom_vline( xintercept = median(SumStepsPerDay), col = "magenta", size = 2)
print(p)
png( file = "./figures/Steps_per_Day_Histogram.png", height = 480, width = 480, unit = "px")
print(p)
dev.off()
```

## Calculates Mean and median of steps taken each day
MeanStepsPerDay <- mean(SumStepsPerDay)
#print(MeanStepsPerDay)
MedianStepsPerDay <- median(SumStepsPerDay)
#print(MedianStepsPerDay)
```

The mean and median of steps taken per day are `r MeanStepsPerDay` and `r MedianStepsPerDay` respectively.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
